---
title: "Two-Way ANOVA (2B) Data Analysis"
output: html_document
date: "2026-04-28"
---

# Introduction

This study examines whether crop yield differs according to:

- Fertilizer Blend
- Crop Type
- Fertilizer × Crop interaction

## Hypotheses

H01: No main effect of fertilizer  
H02: No main effect of crop  
H03: No interaction effect

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(car)
library(emmeans)
library(effectsize)
library(rstatix)
library(knitr)
```

# Data

```{r}
yield <- c(
123,156,112,100,168,
128,150,174,116,109,
166,178,187,153,195,
151,125,117,155,158,

135,130,176,120,155,
175,132,120,187,184,
140,145,159,131,126,
167,183,142,167,168,

156,180,147,146,193,
186,138,178,176,190,
185,206,188,165,188,
175,173,154,191,169
)

fertilizer <- factor(rep(c("BlendX","BlendY","BlendZ"),
each=20))

crop <- factor(rep(
rep(c("Wheat","Corn","Soy","Rice"),
each=5),3))

data <- data.frame(yield,fertilizer,crop)
```

# Descriptive Statistics

```{r}
data %>%
group_by(fertilizer,crop) %>%
summarise(
Valid=n(),
Mean=mean(yield),
SD=sd(yield),
Skewness=e1071::skewness(yield),
Kurtosis=e1071::kurtosis(yield)
)
```

---

# Assumption Checks

## Assumption 1–3
- Continuous DV ✓
- Two categorical IVs ✓
- Independence assumed ✓

---

## Assumption 4: Outliers

```{r}
data %>%
group_by(fertilizer,crop) %>%
identify_outliers(yield)

ggplot(
data,
aes(
crop,
yield,
fill=fertilizer
)
)+
geom_boxplot(
position=position_dodge(.8),
outlier.colour="red"
)+
geom_jitter(
position=position_jitterdodge(),
alpha=.6
)+
theme_minimal()
```

---

## Assumption 5 Normality
(Like your reference — Shapiro per cell)

```{r}
data %>%
group_by(fertilizer,crop) %>%
shapiro_test(yield)
```

Interpret:
p > .05 = normal

---

## Assumption 6 Homogeneity

```{r}
leveneTest(
yield ~ fertilizer*crop,
data=data
)
```

---

# Two-Way ANOVA (Type III SS)

```{r}
options(contrasts=
c("contr.sum","contr.poly"))

model <- lm(
yield ~ fertilizer*crop,
data=data
)

Anova(
model,
type=3
)
```

---

# Effect Sizes
```{r}
install.packages("rstatix")
library(rstatix)

anova_test(
data=data,
dv=yield,
between=c(fertilizer,crop)
)
```
---

# If Interaction Significant:
# Simple Main Effects
(Like your reference)

```{r}
emmeans(model,
pairwise ~ fertilizer|crop,
adjust="bonferroni")
```

```{r}
emmeans(model,
pairwise ~ crop|fertilizer,
adjust="bonferroni")
```

---

# Post Hoc Comparisons

```{r}
pairs(
emmeans(model,
~ fertilizer*crop),
adjust="tukey"
)
```

---

# Interaction Plot

```{r}
interaction.plot(
data$crop,
data$fertilizer,
data$yield
)
```

or

```{r}
ggplot(
data,
aes(crop,yield,
color=fertilizer,
group=fertilizer)
)+
stat_summary(
fun=mean,
geom="line"
)+
stat_summary(
fun=mean,
geom="point"
)
```

---

# APA Results

A two-way factorial ANOVA was conducted to test the effects of fertilizer blend and crop type on yield.

There was a significant main effect of fertilizer,
F(2,48)=14.14, p<.001, partial η²=.37.

There was a significant main effect of crop,
F(3,48)=3.54, p=.021, partial η²=.18.

There was also a significant fertilizer × crop interaction,
F(6,48)=2.64, p=.028, partial η²=.25,
indicating fertilizer effectiveness depended on crop type.

---

# Conclusion

Reject H01  
Reject H02  
Reject H03